**A framework for effective application of machine learning to microbiome-based classification problems**

### 2020-02-24 submission/manuscript.Rmd
Pat's edits
Close to submitting

Begüm D. Topçuoğlu${^1}$, Nicholas A. Lesniak${^1}$, Mack Ruffin${^3}$, Jenna Wiens${^2}$, Patrick D. Schloss^1$\dagger$^

@@

__The computational efficiency of each ML model.__ We compared the training times of the seven ML models. The training times increased with the complexity of the model and the number of potential hyperparameter combinations. Also, the linear models trained faster than non-linear models [Figures S2-S3; Figure 5].

@@

\captionsetup{labelformat=empty}
\captionof{table}{\textbf{Table 1.} Characteristics of the machine learning models in our comparative study.}
\small
\begin{tabular}{|l|l|l|}
\hline

@@

\makecell[l]{Logistic \\regression} & \makecell[l]{A predictive regression analysis when the dependent \\variable is binary.} & Linear \\ \hline

\makecell[l]{SVM with \\linear kernel} & \makecell[l]{A classifier that is defined by an optimal linear \\separating hyperplane that discriminates between labels.} & Linear  \\ \hline

\makecell[l]{SVM with \\radial basis kernel} & \makecell[l]{A classifier that is defined by an optimal non-linear \\separating hyperplane that discriminates between labels.} & Non-linear \\ \hline

\makecell[l]{Decision tree} & \makecell[l]{A classifier that sorts samples down from the
root to the \\leaf node where an attribute is tested to discriminate \\between labels.} & Non-linear \\ \hline

Random forest & \makecell[l]{A classifier that is an ensemble of decision trees \\ that grows randomly with subsampled data.} & Non-linear \\ \hline

\makecell[l]{Gradient Boosted Trees \\ (XGBoost)} & \makecell[l]{A classifier that is an ensembe of decision trees \\ that grows greedily.} & Non-linear \\ \hline

>>>>

Begüm D. Topçuoğlu${^1}$, Nicholas A. Lesniak${^1}$, Mack Ruffin${^3}$, Jenna Wiens^2$\dagger$^, Patrick D. Schloss^1$\dagger$^

@@

__The computational efficiency of each ML model.__ We compared the training times of the seven ML models. The training times increased with the complexity of the model and the number of potential hyperparameter combinations. Also, the linear models trained faster than non-linear models [Figure 5].

@@

**Table 1.** Characteristics of the machine learning models in our comparative study.

\small
\begin{tabular}{|l|l|l|}
\hline

@@

\makecell[l]{Logistic regression} & \makecell[l]{A predictive regression analysis when the dependent variable is binary.} & Linear \\ \hline

\makecell[l]{SVM with linear kernel} & \makecell[l]{A classifier that is defined by an optimal linear separating hyperplane that discriminates between labels.} & Linear  \\ \hline

\makecell[l]{SVM with radial basis kernel} & \makecell[l]{A classifier that is defined by an optimal non-linear separating hyperplane that discriminates between labels.} & Non-linear \\ \hline

\makecell[l]{Decision tree} & \makecell[l]{A classifier that sorts samples down from the
root to the leaf node where an attribute is tested to discriminate between labels.} & Non-linear \\ \hline

Random forest & \makecell[l]{A classifier that is an ensemble of decision trees that grows randomly with subsampled data.} & Non-linear \\ \hline

\makecell[l]{Gradient Boosted Trees (XGBoost)} & \makecell[l]{A classifier that is an ensembe of decision trees that grows greedily.} & Non-linear \\ \hline
